---
title: "Argument Passing"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Argument Passing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

A big change to the {EGAnet} structure in version +2.0.0 was that nearly every argument can be passed down through sub-functions in a main function. Sounds useful, but what does that actually mean?

Let's take the `EGA` function as an example. The `EGA` function has several other {EGAnet} functions that it uses to perform it's operations. These functions include:

+   `auto.correlate`
    -   `cor`
    -   `polychoric.matrix`

+   `EGA.estimate`
    -   `auto.correlate` (see functions above)
    -   `network.estimation`
        +   `BGGM`
        +   `EBICglasso.qgraph`
        +   `TMFG`
    -   `community.detection` (see `igraph::cluster_*` functions)
    -   `community.consensus`
    
+   `community.unidimensional`
    -   `auto.correlate` (see functions above)
    -   `network.estimation` (see functions above)
    
+   `plot`

The first-level points are all functions called by `EGA`. The second-level points are called by their first level functions and so on down the line. This function tree shows that (1) `EGA` calls a lot of functions but (2) all of these functions have their own arguments that *can* be called by `EGA`.

`EGA` passes arguments that are not listed as default arguments (see `?EGA`) using the `...` argument. For every function in {EGAnet} that has the `...` means that you can pass along arguments to other functions in the package (and other packages). How do you know what arguments can be passed? The documentation for the `...` argument in every function with tell you. Let's take a look at `EGA`'s `...` argument documentation:

> 	Additional arguments to be passed on to `auto.correlate`, `network.estimation`, `community.detection`, `community.consensus`, and `community.unidimensional`

Each of these functions are linked directly to their documentation, so you can navigate to each function when specifying arguments of an {EGAnet} function.

OK, let's see this flexibility in action.

### Passing through `auto.correlate`

#### Changing `corr`

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
# Load packages
library(EGAnet); library(psychTools)

# Use Spearman correlation
bfi_ega_spearman <- EGA(bfi[,1:25], corr = "spearman", plot.EGA = FALSE)

# Print result
summary(bfi_ega_spearman)
```

How do you know whether something has changed? In general, most changes to methods will show up in the summary of the output. In this example, we can see that `Correlations: spearman` meaning that the correlations were changed to Spearman.

#### Changing `ordinal.categories`

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
# 5 categories and higher are treated as continuous
bfi_ega_continuous <- EGA(bfi[,1:25], ordinal.categories = 4, plot.EGA = FALSE)

# Print result
summary(bfi_ega_continuous)
```

There usually isn't a case where you would want to change `ordinal.categories` argument of `auto.correlate` to be lower than its default value of `7`. The argument sets what data the function should consider as ordinal (defaults to up to 7 categories). For demonstration purposes, `ordinal.categories` is used to show that arguments from `EGA` are passed on to `auto.correlate`.

### Passing through `network.estimation`

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
# Change the lambda.min.ratio
bfi_ega_qgraph <- EGA(bfi[,1:25], lambda.min.ratio = 0.01, plot.EGA = FALSE)

# Print result
summary(bfi_ega_qgraph)
```

The `lambda.min.ratio` parameter was passed to `network.estimation` and on to `EBICglasso.qgraph`. The parameter is output in the summary above (see `Lambda:`). The `lambda.min.ratio` value used in this example is the default in `qgraph::EBICglasso`.

### Passing through `community.detection`

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
# Change the algorithm and its arguments
bfi_ega_leiden <- EGA(
  bfi[,1:25], algorithm = "leiden",
  objective_function = "CPM", # in {igraph}
  resolution_parameter = 0.05, # in {igraph}
  plot.EGA = FALSE
)

# Print result
summary(bfi_ega_leiden)
```

The above code shows off *a lot* of {EGAnet}'s flexibility. First, the algorithm is being changed to `"leiden"` which is switched in the `community.detection` function. Second, the arguments `objective_function` and `resolution_parameter` are being passed to `igraph::cluster_leiden` and aren't even handled in {EGAnet}. The summary reflects these changes: `Algorithm: Leiden with Constant Potts Model`. 

### Passing through `community.consensus`

#### Changing `consensus.iter`

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
# More consensus iterations
bfi_ega_consensus <- EGA(
  bfi[,1:25], algorithm = "louvain",
  consensus.iter = 10000, plot.EGA = FALSE
)

# Print result
summary(bfi_ega_consensus)
```

By default, the consensus clustering method performs `1000` iterations (see `?consensus.clustering`) but the iterations can be changed as well as the method (using `consensus.method` argument). This change is also reflected in the summary output with number of iterations in parentheses.

#### Changing `order`

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
# Get lower order result
bfi_ega_lower <- EGA(
  bfi[,1:25], algorithm = "louvain",
  order = "lower", plot.EGA = FALSE
)

# Print result
summary(bfi_ega_lower)
```

The `order` can also be changed. The `"lower"` order is used in Hierarchical EGA (`hierEGA`; [Jiménez et al., 2023](https://doi.org/10.1037/met0000590)) method but can be replicated using `EGA` and specifying the `order` argument.  