---
title: "EGA + Transformers NLP"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{EGA + Transformers NLP}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Transformers models are taking the data science world by storm. Their rise in popularity is due to their unparalleled ability to understand contextual information in language. In short, these models can quantify the qualitative.

To demonstrate how transformers natural language processing (NLP) can be used in combination with {EGAnet}, we'll use an example of Taylor Swift's first and most recent albums: *Taylor Swift* and *Midnights*.

### Set Up Transformers NLP and Genius API

To get started, the [{text}](https://r-text.org) package needs to be installed and set up (see [{text} installation](http://r-text.org/articles/huggingface_in_r_extended_installation_guide.html) for instructions on getting set up).

Once set up, load {text}:

```{r, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
load("./data/ega-nlp.RData")
```

```{r, echo = TRUE, eval = FALSE, warning = FALSE, message = FALSE, comment = NA}
library(text)
```

```{r, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
EGAnet:::silent_load(library(text))
```

We'll use another package called {simpleRgenius} to scrape lyrics from the [Genius](https://genius.com/) lyrics website. This package needs to be installed from GitHub and loaded:

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
# Install {devtools} (if necessary)
if(!"devtools" %in% unlist(lapply(.libPaths(), list.files))){
  install.packages("devtools")
}

# Install {simpleRgenius}
devtools::install_github("AlexChristensen/simpleRgenius")

# Load {simpleRgenius}
library(simpleRgenius)
```

After installing {simpleRgenius}, you'll need to set up an API with Genius (only necessary to reproduce this example). You can follow the instructions [here](https://github.com/AlexChristensen/simpleRgenius/wiki/Genius-API-Setup).

### Scraping Lyrics

At this point, you should have loaded {text} and {simpleRgenius} as well as imported your Genius API into R's environment (if following along and reproducing the code). The next step is to get all of the songs for Taylor Swift's *Midnights* album:

```{r, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
Sys.setenv(GENIUS_API_TOKEN = "xT4ny-9oP9kQXKmG1NtTRcLfzaHIc02R-fWEM_bbU2Bp_uHK-zHzVs4929vWR0Pz")
```

```{r, echo = TRUE, eval = FALSE, warning = FALSE, message = FALSE, comment = NA}
# First, let's get all of the song names
midnights_songs <- c(
  "Lavender Haze", "Maroon", "Anti-Hero",
  "Snow On The Beach", "You're On Your Own, Kid",
  "Midnight Rain", "Question...?", "Vigilante Shit",
  "Bejeweled", "Labyrinth", "Karma",
  "Sweet Nothing", "Mastermind", 
  "The Great War", "Bigger Than The Whole Sky",
  "Paris", "High Infidelity", "Glitch",
  "Would've, Could've, Should've",
  "Dear Reader"
)

# Next, let's get the lyrics
midnights_lyrics <- get_lyrics(
  artist_name = "Taylor Swift",
  song_names = midnights_songs
)
```

### Perform Zero-shot Classification

Next, zero-shot classification can be performed. There are many different models on [huggingface](https://huggingface.co) to choose from. We'll use Cross Encoder's base [RoBERTa](https://huggingface.co/cross-encoder/nli-roberta-base) model:

```{r, echo = TRUE, eval = FALSE, warning = FALSE, message = FALSE, comment = NA}
midnights_zero <- textZeroShot(
  sequences = midnights_lyrics$Lyric, # text
  candidate_labels = c(
    "anxiety", "depression", "happiness",
    "heartbreak", "love", "relationships",
    "rebellious", "revenge", "romance"
  ),
  # theme (same as sentiment analysis)
  model = "cross-encoder/nli-roberta-base",
  # model to use
  multi_label = TRUE
  # whether multiple labels can be true
)
```

### Reformat Zero-shot Output into Long Format

To analyze data with {EGAnet}, data need to be in wide format. But first, let's do some basic NLP and visualization. The following code will (1) reformat the data into long format, (2) remove non-existent labels, (3) add back song names, and (4) summarize the lyrics by song:

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
# Load {tidyverse}
library(tidyverse)

# Wrangle the data
midnights_long <- midnights_zero %>%
  pivot_longer(
    cols = starts_with("labels_") | starts_with("scores_"),
    # Obtain columns with labels and scores
    names_to = c(".value", "label_number"),
    # Push the names to new columns
    names_pattern = "(labels|scores)_(x_.*)"
    # Set up the patterning for labels and scores
  )

# Remove NAs
midnights_long <- na.omit(midnights_long)

# Initialize vector for song names
song_names <- character(length = nrow(midnights_long))

# Create loop to store song names
for(i in 1:nrow(midnights_long)){
  
  # Get matches for lyric
  lyric_match <- min(
    match( # Use `min` to get only one index
      midnights_long$sequence[i],
      midnights_lyrics$Lyric
    )
  )
  
  # Get song name
  song_names[i] <- midnights_lyrics$Song[lyric_match]
  
}

# Create vector to add songs back
midnights_long$song_name <- song_names

# Summarize by song
midnights_summary <- midnights_long %>%
  group_by(song_name, labels) %>%
  summarize(
    Probability = mean(
      scores, na.rm = TRUE
    )
  )
```

### Basic NLP Visualization

#### Keep Order of Songs

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
# Before visualizing...
## Create factor to keep songs in order
midnights_summary$song_name <- factor(
  midnights_summary$song_name,
  levels = midnights_songs
)
```

#### Bar Plot by Song

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA, fig.height = 8, fig.width = 13, fig.align = "center"}
# Visualize using histogram
ggplot(
  data = midnights_summary,
  aes(x = Probability, y = labels, fill = labels)
) +
  facet_wrap(~song_name) +
  geom_histogram(stat = "identity") +
  labs(
    x = "Probability",
    y = "Sentiment",
    title = "Midnights Album Sentiment"
  ) +
  scale_y_discrete(limits = rev) + # reverse order
  theme( # basic theme
    panel.background = element_blank(),
    legend.title = element_blank(),
    plot.title = element_text(size = 16, hjust = 0.5),
    axis.line = element_line(linewidth = 1),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    strip.text = element_text(size = 10),
    legend.text = element_text(size = 10),
    axis.text.x = element_text(angle = 45, hjust = 1)
  )
```

#### Album Arch of "relationships" Theme

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA, fig.height = 6, fig.width = 10, fig.align = "center"}
# Visualize using line plot
midnights_summary %>%
  mutate( # focus on relationships across songs
    alpha = ifelse(labels == "relationships", 1, 0.25)
  ) %>%
  ggplot(
    aes(
      x = song_name, y = Probability, group = labels, 
      color = labels, alpha = alpha
    ) 
  ) +
  geom_line(linewidth = 1) +
  labs(
    x = "Song",
    y = "Probability",
    title = "Midnights Album Arch"
  ) +
  theme( # basic theme
    panel.background = element_blank(),
    legend.title = element_blank(),
    plot.title = element_text(size = 16, hjust = 0.5),
    axis.line = element_line(linewidth = 1),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 14),
    legend.text = element_text(size = 10),
    legend.position = "bottom",
    legend.key = element_rect(fill = NA), # remove grey box
    axis.text.x = element_text(angle = 45, hjust = 1)
  ) +
  scale_alpha_identity(guide = "none")
```

### Get Dimensionality of Album

#### Reformat Lyrics into Wide Format

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
# Make each sentiment a column with values
midnights_wide <- midnights_long %>%
  pivot_wider(
    names_from = "labels",
    values_from = "scores"
  ) %>%
  group_by(song_name, sequence) %>%
  summarize( 
    anxiety = sum(anxiety, na.rm = TRUE), # CHANGE!
    depression = sum(depression, na.rm = TRUE),
    happiness = sum(happiness, na.rm = TRUE),
    heartbreak = sum(heartbreak, na.rm = TRUE),
    love = sum(love, na.rm = TRUE),
    relationships = sum(relationships, na.rm = TRUE),
    rebellious = sum(rebellious, na.rm = TRUE),
    revenge = sum(revenge, na.rm = TRUE),
    romance = sum(romance, na.rm = TRUE)
  )
```

#### Perform EGA

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
# Load {EGAnet}
library(EGAnet)

# Perform EGA
midnights_ega <- EGA(midnights_wide[,-c(1:2)])

# Get summary
summary(midnights_ega)
```

#### Check Stability of EGA

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
# Perform bootstrap EGA
midnights_boot <- bootEGA(midnights_wide[,-c(1:2)], seed = 12)

# Get summary
summary(midnights_boot)

# Perform dimension stability
dimensionStability(midnights_boot)
```

Overall, the dimensions are relative stable. The `depression` theme goes between the first and second dimension (roughly positive and negative valence themes, respectively). 

### Invariance between the *Taylor Swift* and *Midnights* Albums

Taking this analysis a step further, the themes can be analyzed for configural and metric invariance to determine whether there are thematic differences in lyrics between Taylor Swift's first and most recent albums.

This analysis mimics analyses that might be performed using text data from an intervention or therapy study (e.g., pre/post, experiment/control).

The same zero-shot classification analyses (including the same Cross Encoder RoBERTa model) should be performed using the same themes (i.e., classes).

```{r, echo = FALSE, eval = FALSE, warning = FALSE, message = FALSE, comment = NA}
# Get song names
swift_songs <- c(
  "Tim McGraw", "Picture to Burn", "Teardrops On My Guitar",
  "A Place In This World", "Cold as You", "The Outside",
  "Tied Together with a Smile", "Stay Beautiful",
  "Should've Said No", "Mary's Song (Oh My My My)",
  "Our Song", "I'm Only Me When I'm With You", "Invisible"
)

# Get lyrics
swift_lyrics <- get_lyrics(
  artist_name = "Taylor Swift",
  song_names = swift_songs
)

# Perform zero-shot classification
swift_zero <- textZeroShot(
  sequences = swift_lyrics$Lyric, # text
  candidate_labels = c( # themes
    "anxiety", "depression", "happiness",
    "heartbreak", "love", "relationships",
    "rebellious", "revenge", "romance"
  ),
  model = "cross-encoder/nli-roberta-base",
  multi_label = TRUE
)

# Wrangle the data
swift_long <- swift_zero %>%
  pivot_longer(
    cols = starts_with("labels_") | starts_with("scores_"),
    # Obtain columns with labels and scores
    names_to = c(".value", "label_number"),
    # Push the names to new columns
    names_pattern = "(labels|scores)_(x_.*)"
    # Set up the patterning for labels and scores
  )

# Remove NAs
swift_long <- na.omit(swift_long)

# Initialize vector for song names
song_names <- character(length = nrow(swift_long))

# Create loop to store song names
for(i in 1:nrow(swift_long)){
  
  # Get matches for lyric
  lyric_match <- min(
    match( # Use `min` to get only one index
      swift_long$sequence[i],
      swift_lyrics$Lyric
    )
  )
  
  # Get song name
  song_names[i] <- swift_lyrics$Song[lyric_match]
  
}

# Create vector to add songs back
swift_long$song_name <- song_names

# Summarize by song
swift_summary <- swift_long %>%
  group_by(song_name, labels) %>%
  summarize(
    Probability = mean(
      scores, na.rm = TRUE
    )
  )

# Before visualizing...
## Create factor to keep songs in order
swift_summary$song_name <- factor(
  swift_summary$song_name,
  levels = swift_songs
)

# Make each sentiment a column with values
swift_wide <- swift_long %>%
  pivot_wider(
    names_from = "labels",
    values_from = "scores"
  ) %>%
  group_by(song_name, sequence) %>%
  summarize( 
    anxiety = sum(anxiety, na.rm = TRUE), # CHANGE!
    depression = sum(depression, na.rm = TRUE),
    happiness = sum(happiness, na.rm = TRUE),
    heartbreak = sum(heartbreak, na.rm = TRUE),
    love = sum(love, na.rm = TRUE),
    relationships = sum(relationships, na.rm = TRUE),
    rebellious = sum(rebellious, na.rm = TRUE),
    revenge = sum(revenge, na.rm = TRUE),
    romance = sum(romance, na.rm = TRUE)
  )

# Perform EGA
swift_ega <- EGA(swift_wide[,-c(1:2)])

# Get summary
summary(swift_ega)

# Perform bootstrap EGA
swift_boot <- bootEGA(swift_wide[,-c(1:2)], seed = 12)

# Get summary
summary(swift_boot)

# Perform dimension stability
dimensionStability(swift_boot)
```

#### Set Up for Invariance

Check out the mean probabilities of each theme across the albums

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
# Look at the mean probabilities of both
# the Taylor Swift and Midnights albums
colMeans(swift_wide[,-c(1:2)])
colMeans(midnights_wide[,-c(1:2)])
```

Get *p*-values for *t*-tests between each theme

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
sapply(
  colnames(swift_wide)[-c(1:2)], function(theme){
    t.test(
      swift_wide[,theme], midnights_wide[,theme], 
      var.equal = TRUE
    )$p.value
  }
)
```

Based on means, there are differences between `happiness` and `rebellious`.

To get set up for `invariance`, groups need to be created:

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
# Set up groups
groups <- c(
  rep("Taylor Swift", nrow(swift_wide)),
  rep("Midnights", nrow(midnights_wide))
)

# Combine datasets
combined_wide <- rbind(
  swift_wide[,-c(1:2)], midnights_wide[,-c(1,2)]
)
```

#### Perform Invariance

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA, fig.height = 6, fig.width = 8, fig.align = "center"}
# Perform invariance
ega_invariance <- invariance(
  data = combined_wide,
  groups = groups,
  seed = 12
)

# View and plot results
ega_invariance; plot(ega_invariance)
```

Based on `invariance`, there is configural invariance but metric non-invariance for `anxiety`, `depression`, and `rebellious` themes. `anxiety` and `depression` themes had greater connectivity in the *Taylor Swift* album while `rebellious` themes had greater connectivity in the *Midnights* album. In essence, the two albums captured relatively different negative themes.