---
title: "Errors"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Errors}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

For every function in {EGAnet}, there are built-in data checks that may sometimes throw an error. Errors in R can be cryptic and {EGAnet} tries to help out by having some systematic errors that are a bit more user-friendly.

This guide is aimed at making the data check errors that {EGAnet} throws more transparent. Below are some common data check errors. If you find one of these errors, then please **do not** submit a bug report -- these errors are by design! Try to fix whatever the error is telling you as it is likely something quick that can be handled in your data.

### "range" Error

A range error is when a function receives data (numeric values) that are outside of the expected range of values. This error most frequently occurs with `polychoric.matrix` in `auto.correlate` because of the underlying C code.

To set up an example, we'll recode the optimism dataset to be on a 5-point Likert scale from -2 to 2 rather than 1 to 5:

```{r, echo = TRUE, eval = FALSE, warning = FALSE, message = FALSE, comment = NA}
# Load {EGAnet}
library(EGAnet)

# Recode the optimism dataset
optimism_recoded <- apply(optimism, 2, function(x){x - 3})

# Estimate EGA
ega_optimism <- EGA(data = optimism_recoded)
```

```{r, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
# Load {EGAnet}
library(EGAnet)

# Recode the optimism dataset
optimism_recoded <- apply(optimism, 2, function(x){x - 3})
```

```{r, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
cat("Error in \"polychoric.matrix\" :
Minimum of 'data' (-2) does not match expected range(s). Range must be between: '0' and '11'")
```

The error specifies that the issue occurred in the `polychoric.matrix` function. The error says that there are data with a minimum value of `-2` in the dataset when the function expects a minimum value to be `0`.

Without getting bogged down in the details, the `polychoric.matrix` function is fast and efficient for computing polychoric correlations, but in order to make it so, it needs to take advantage of clever indexing. In C, indices start at `0`, so any values lower than `0` means that the data need to be re-indexed so that the lowest value is `0` (and the highest value, if applicable, is `11`).

To fix this issue, you can recode your data:

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
# Get the minimum value
minimum_value <- min(optimism_recoded, na.rm = TRUE)

# Print value
minimum_value

# Recode the values to start at zero
optimism_fixed <- apply(
  optimism_recoded, 2, function(x){
    x + abs(minimum_value)
  }
)

# Estimate EGA
ega_optimism <- EGA(data = optimism_fixed)
```

By recoding the lowest value to `0`, the error goes away and `EGA` executes successfully.

### "class" Error

A class error is when a function receives a class it does not expect:

```{r, echo = TRUE, eval = FALSE, warning = FALSE, message = FALSE, comment = NA}
# Estimate EGA
ega_optimism <- EGA(data = optimism)

# Compute the Generalized Total Entropy Fit Index
genTEFI(ega_optimism)
```

```{r, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
cat("Error in \"genTEFI\" : 
  Input into 'data' is not the expected class 'hierEGA'. Input is class 'EGA'")
```

The `genTEFI` function expects a `hierEGA` class but it received an `EGA` class object. This error means that the wrong object class was used with the function.

### "object" Error

A object error is when a function receives a object that it is not designed to handle:

```{r, echo = TRUE, eval = FALSE, warning = FALSE, message = FALSE, comment = NA}
# Estimate EGA using a list object
opt.hier <- EGA(data = as.list(optimism))
```

```{r, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
cat("Error in \"EGA\" :
Input into 'data' argument is not 'matrix', 'data.frame', 'tibble'. Input is 'list'")
```

Here, `EGA` expects to receive a matrix, data frame, or tibble but the input was a list. The data will need to be converted to one of these three object types before proceeding with the analysis.

### "typeof" Error

A typeof error is when a function receives data that is not in the expected type:

```{r, echo = TRUE, eval = FALSE, warning = FALSE, message = FALSE, comment = NA}
# Estimate EGA using a character TRUE
opt.hier <- EGA(data = optimism, plot.EGA = "TRUE")
```

```{r, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
cat("Error in \"EGA\" :
Input into 'plot.EGA' is 'character'. Input is expected to be 'logical'")
```

In this example, `EGA` expected input into the argument `plot.EGA` to be a logical (`TRUE` or `FALSE`) but instead got a character `"TRUE"`. The error tells you that the input is expected to be logical.

### "length" Error

A length error is when a function receives input that is longer (or shorter) than expected:

```{r, echo = TRUE, eval = FALSE, warning = FALSE, message = FALSE, comment = NA}
# Estimate EGA using a character TRUE
opt.hier <- EGA(data = optimism, plot.EGA = c(TRUE, TRUE))
```

```{r, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
cat("Error in \"EGA\" :
Length of 'plot.EGA' (2) does not match expected length(s). Length must be: '1'")
```

This time for the `plot.EGA` argument the correct type of input was used (logical) but there were too many input. The `(2)` says that the argument received two inputs when it expected only `1`.

### None of the Above

If you've made it this far and none of the above errors apply, then submit an [issue on GitHub](https://github.com/hfgolino/EGAnet/issues). 

*Make sure to provide a minimal reproducible example (like those demonstrated above) and send any data necessary to expedite the process of getting the error fixed.*