---
title: "Using External Networks (and external packages)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using External Networks (and external packages)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Although {EGAnet} features many common methods used in the network psychometric literature, it does not include all possible options. Sometimes you might switch between {EGAnet} and other packages.

### Alternative Networks

There are many different packages available to estimate psychometric networks. One common package is [{bootnet}](https://cran.r-project.org/web/packages/bootnet/index.html). `bootnet::estimateNetwork` offers many different methods to estimate networks. The `"ggmModSelect"` approach will be used as an example.

```{r, echo = TRUE, eval = FALSE, warning = FALSE, message = FALSE, comment = NA}
# Load packages
library(EGAnet); library(bootnet)

# Load data
data <- wmt2[,7:24]

# Estimate network
stepwise_result <- estimateNetwork(
  data = data, default = "ggmModSelect", stepwise = TRUE
)
```

```{r, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
# Load packages
library(EGAnet); library(bootnet)

# Load data
data <- wmt2[,7:24]

# Estimate network
sink <- capture.output(
  stepwise_result <- estimateNetwork(
    data = data, default = "ggmModSelect", stepwise = TRUE
  )
)
```

Once a network is estimated, then a modular approach to estimating the `EGA` workflow can be used. This workflow is the same as what's used internally in the `EGA` function. If the network is coming from an [{igraph}](https://r.igraph.org/) must be converted using the `igraph2matrix` function (e.g., `ega_network <- igraph2matrix(igraph_network)`). Any network that is a matrix or data frame object can be used in {EGAnet} functions like the output from `estimateNetwork`.

#### Check for Unidimensionality

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
# Perform unidimensionality check
unidimensional_membership <- community.unidimensional(data)

# Print
unidimensional_membership
```

The standard unidimensional check uses the Louvain [Louvain](https://doi.org/10.1088/1742-5468/2008/10/P10008) algorithm on the zero-order correlation matrix ([Christensen, 2023](https://doi.org/10.31234/osf.io/ep3vx)). The output will include the memberships regardless of whether the data are detected as unidimensional. In this output, the number of communities is 2 and therefore *not* unidimensional. If the number of communities was 1, then there is no need to proceed with the multidimensional estimation.

#### Estimate Multidimensionality

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
# Estimate multidimensionality
multidimensional_membership <- community.detection(
  stepwise_result$graph, algorithm = "walktrap"
)

# Print
multidimensional_membership
```

For the multidimensional estimation, the estimated network should be used as the input. The algorithm can be set using a number of different algorithms (see `?community.detection`) but the default is to use the [Walktrap](https://www.emis.de/journals/JGAA/getPaper-437.html?id=30) algorithm. From this output, Walktrap estimates 2 communities.

#### Obtain Final Memberships

A shortcut to obtain the final memberships as is used in `EGA` is provided below:

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
wc <- EGAnet:::swiftelse(
    # Check for whether unidimensional membership should be used
    EGAnet:::unique_length(unidimensional_membership) == 1,
    unidimensional_membership, multidimensional_membership
    # Otherwise, use multidimensional membership
  )
```

#### Plot

In order to plot using {EGAnet}, the network and memberships need to be set up as an `EGA` class object:

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
# Set up EGA object
ega_object <- list(
  network = stepwise_result$graph,
  wc = wc
)

# Set `EGA` class
class(ega_object) <- "EGA"

# Plot
plot(ega_object)
```

#### Modularity

Modularity can also be computed using the estimated network and appropriate memberships.

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
modularity(stepwise_result$graph, wc)
```

### Alternative Similarity Measures

By default, {EGAnet} uses the `auto.correlate` function to compute appropriate correlations for each set of pairwise variables. Other measures might need to be used in {EGAnet} functions. 

#### Cosine Similarity

A common example is from natural language processing and cognitive science where cosine similarity is a common association measure between two terms. Below, we'll show a semantic network example from the [{SemNeT}](https://doi.org/10.1037/met0000463) package.

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
# Load packages
library(EGAnet); library(SemNeT)

# Compute cosine similarity
animals_cosine <- similarity(open.binary, method = "cosine")
```

This data are from participants who performed a verbal fluency task and generated animals for 1 minute. Each row represents a participant and each column represents an animal. A `1` is an animal a participant provided; a `0` is an animal a participant did not provide.

The `animals_cosine` object is a symmetric matrix that represents the cosine similarity between each animal. To use the cosine similarity matrix (or any other alternative similarity matrix) in an {EGAnet} function, `n` or the number of cases must be set:

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA, fig.height = 6, fig.width = 7, fig.align = "center"}
# Compute EGA
animals_ega <- EGA(
  # Arguments for `EGA`
  animals_cosine, n = nrow(open.binary), model = "TMFG",
  # Arguments for `plot`
  node.size = 4, label.size = 3, edge.size = 2
)
```