---
title: "Plotting with {EGAnet}"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Plotting with {EGAnet}}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Plotting in {EGAnet} has become much more flexible by passing on _most_ arguments to {GGally}'s [ggnet2](https://briatte.github.io/ggnet/) (some arguments might not change to keep a consistent {EGAnet} style to the plots). This Wiki walks through some of the flexibility available in {EGAnet}'s plots. This demonstration is only the start &ndash; you can use these plots as the foundation for your own creations. These examples focus solely on `EGA` plotting but apply to all `*EGA` plots. 

### Basic Plot

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
# Load necessary packages
library(EGAnet); library(GGally); library(ggplot2)

plot(ega.wmt)
```

### Legend

#### Change Names

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
plot(ega.wmt, legend.names = c("Dimension 1", "Dimension 2"))
```

#### Remove legend

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
plot(ega.wmt) + theme(legend.position = "none")
```

Handling legend title and text can be done using the standard {ggplot2} `theme` arguments

### Title

#### Basic Title

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
plot(ega.wmt) + ggtitle("EGA Title")
```

#### Adjusted Title (centered, larger size, bold)

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
plot(ega.wmt) + 
  ggtitle("Better EGA Title") +
  theme(plot.title = element_text(size = 14, face = "bold", hjust = 0.5))
```

### Layout

For all available layout options, see documentation for {sna}'s `gplot.layout`.

#### Circle

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
plot(ega.wmt, layout = "circle")
```

#### Eigen

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
plot(ega.wmt, layout = "eigen")
```

### Nodes

#### Colors

Change Colors Using Color Palette

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
plot(ega.wmt, color.palette = "blue.ridge2")
```

Change Colors Using HEX Codes

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
plot(ega.wmt, node.color = c("#F5815A", "#47BCC9"))
```

An awesome resource to generate color palettes: [https://coolors.co/](https://coolors.co/)

#### Size

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
plot(ega.wmt, node.size = 12)
```

Change Size based on Node Strength

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
plot(ega.wmt, node.size = colSums(ega.wmt$network)^2 * 16)
```

### Edges

#### Size

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
plot(ega.wmt, edge.size = 12)
```

#### Transparency

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
plot(ega.wmt, edge.alpha = 0.1)
```

#### Colors and Line Type

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
plot(
  ega.wmt, edge.alpha = 1,
  edge.lty = c( # line type
    "dashed", # positive edge first
    "solid" # negative edge second
  ),
  edge.color = c( # color
    "pink", # positive edges first
    "black" # negative edges second
  )
)
```

#### Positive Only

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
plot(
  ega.wmt,
  edge.color = c(
    "darkgreen", # positive edges (default color)
    "white" # negative edges second
  )
)
```

#### Negative Only

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
plot(
  ega.wmt, edge.alpha = 1, # used to increase visibility 
  edge.color = c(
    "white", # positive edges first
    "red" # negative edges (default color)
  )
)
```