---
title: "Quick Start"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick Start}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

To get started with {EGAnet}, you can use the Wiener Matrizen Test 2 (WMT-2) that is included in the package. The WMT-2 dataset is a German matrices intelligence test (in English, the name translates to Vienna Matrices Test 2). The test was designed using Rasch modeling to derive a short measure of (fluid) intelligence.

The dataset is comprised of some demographic variables as well as the test items. Here's the variable names:

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
# Load {EGAnet}
library(EGAnet)

# Check variables names
colnames(wmt2)
```

For the EGA analysis, only the variables of interest are necessary -- that is, only the `"wmt*"` variables are necessary. EGA is performed on the subset of these variables below:

```{r, echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
# Perform EGA
ega_wmt <- EGA(wmt2[,7:24])
```

That's it! You're up and running with `EGA`.

To explore other features of the package, some quick links are provided below:

+   [EGA Workflow](ega.html)

+   [Argument Passing](argument-flexibility.html)

+   [EGA + Transformers NLP](ega-nlp.html)

+   [External Networks (and external packages)](modular-approach.html)

+   [Plotting](plotting.html)