---
title: "What's Changed?"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{What's Changed?}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

In a word: *everything*

Nearly **every** line of code is new. 

New, however, doesn't mean throwing out the old &ndash; your previous code should still work. 

So, what's changed?

### Faster

The new {EGAnet} is *fast*. Nearly all functions were optimized, resulting in 2-15*x* *faster* code than previous versions. Analyses that used to take minutes now take seconds. 

The biggest difference? Fast computation of tetrachoric and polychoric correlations

These correlations are at the core of most {EGAnet} functions. Unfortunately, there aren't many implementations in R that quickly compute these correlations. So, we've implemented our own tetrachoric and polychoric correlations in C. 

Here's a comparison between {EGAnet}'s `auto.correlate`, its former implementation of {qgraph}'s `cor_auto`, {psych}'s `polychoric`, and recently [published](https://doi.org/10.1080/10705511.2021.1929996) and fast implementation in R: {Turbofuns}'s `PolychoricRM` (using 1 core). This demonstration used the `bfi` dataset in {psych} with 100 repetitions:

```{r, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
cat("Unit: milliseconds
      expr       min        lq       mean     median         uq       max neval cld
    EGAnet   75.1616   75.4386   77.57433   75.74405   78.19075  201.0211   500  a 
 Turbofuns   79.5288   79.7113   80.45382   79.87415   80.30210   97.3525   500  a 
    qgraph  437.7368  449.0530  490.07021  456.72130  560.37485  743.6972   500   b 
     psych 2555.8532 2586.6863 2701.75927 2603.60790 2727.74910 4682.7300   500   c")
```

`auto.correlate`'s median time was on par with `PolychoricRM`. `auto.correlate` includes several checks to compute Pearson's correlations for continuous data and polyserial correlations for continuous with ordinal data similar to `cor_auto`; however, `auto.correlate` is *6x faster*. Compared to `polychoric`, `auto.correlate` is *33x faster*.

To put the speed difference in context: `auto.correlate` completes one full `bootEGA` and gets 100 iterations into a *second* `bootEGA` by the time `cor_auto` completes its first 100 iterations in its *first* `bootEGA`.

### More Flexible

Functions are more modular and build on each other to allow for combinations that were previously unthinkable.

Want to use {igraph}'s `cluster_spinglass` in `EGA`? You can do that

Want to see `itemStability` of your `hierEGA`? You can do that

Want to perform a `bootEGA` with `EGA.fit` using `BGGM` and consensus clustering Louvain (`community.consensus`)? That's a lot... but you can do that too 

Such flexibility warrants a disclaimer: Just because you *can*, doesn't mean you *should!* Keep best practices in mind.

### More Informative

Functions that had S3 `print`, `summary`, and `plot` methods have been updated. Functions that didn't have `print`, `summary`, and `plot` methods now do. 

`print` and `summary` methods are designed to give you the information you want without hassle. Here's `EGA`'s `summary` output:

```{r, echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA}
cat("Model: GLASSO (EBIC with gamma = 0.5)
Correlations: auto
Lambda: 0.0764652282008741 (n = 100, ratio = 0.1)

Number of nodes: 25
Number of edges: 117
Edge density: 0.390

Non-zero edge weights: 
     M    SD    Min   Max
 0.046 0.119 -0.269 0.548

----

Algorithm:  Walktrap

Number of communities:  5

A1 A2 A3 A4 A5 C1 C2 C3 C4 C5 E1 E2 E3 E4 E5 N1 N2 N3 N4 N5 O1 O2 O3 O4 O5 
 1  1  1  1  1  2  2  2  2  2  3  3  3  3  3  4  4  4  4  4  5  5  5  5  5 

----

Unidimensional Method: Louvain
Unidimensional: No

----

TEFI: -27.335")
```

Model and algorithm information, network descriptives, community memberships, and fit (Total Entropy Fit Index) are provided with every `*EGA` function. This information is everything (and perhaps more!) than you need to write your reproducible Methods section. 

What about `plot`? There's so much flexibility there that it needs its own [Wiki](https://github.com/hfgolino/EGAnet/wiki/Plots-in-%7BEGAnet%7D)

### More Reproducible

Like polychoric correlations, we've implemented high quality open-source pseudorandom number generation (PRNG) in C instead of R. This change avoids potential conflicts with the `seed` you set in your script and the `seed` you set in an {EGAnet} PRNG function. 

In short, your results can be replicated, time and time again.

*For the gritty details, head over to [Reproducibility and PRNG](https://github.com/hfgolino/EGAnet/wiki/Reproducibility-and-PRNG)*

### For more details on what's changed, check out the [NEWS](https://github.com/hfgolino/EGAnet/blob/master/NEWS)
